Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-01-05 WSSSPE3/report/WSSSPE3_review.tex
removing some line feeds
Small changes

WSSSPE3 left the traditional Supercomputing Conference environment this year,
and in Turk's words, this spoke to the fact that scientific software comes from
many different types of inquiries, deployment, strategies for maintenance,
users, and ways of measuring the value of a piece of software. It appeared to
Turk that the supercomputing community generally adopts some top-down
approaches, whereas the SciPy community more often than not uses more bottom-up
systems. The essential messages perceived were also often bipolar: the
supercomputing community thinks that software is getting harder, with exascale
computing and optimization issues in mind; but the SciPy community thinks that
software is becoming better, with emerging tools such as Jupiter and
productivity packages for research workflows. Admitting such comparisons are
somewhat unfair generalizations, Turk reminded the audience that the different
approaches bring different types of ideas to the table, and he welcomed WSSSPE3
being conducted outside existing preconceptions.

@@

Turk reminded the audience that the volunteer model means unpaid labor. On this
note, he recommended Ashe Dryden's blog post on the ethics of unpaid labor and
the open source software
community\footnote{http://www.ashedryden.com/blog/the-ethics-of-unpaid-labor-and-the-oss-community}.
Often times, a person funded to work full time on a scientific project can spend
a small amount of time for working on a piece of software necessary for that
project. However, researchers' abilities to participate in that volunteer
community are not always the same and may not always be aligned with their
research projects. From Turk's experience, we cannot always rely on unpaid labor
and volunteer time to sustain a piece of software---this came down to the
notions of the top-down and the bottom-up approaches, i.e., the funded versus
the grassroots. However, Turk pointed out that bottom-up, volunteer-driven
projects can be just as large-scale as a top-down software development project.

@@

Work by Katz was supported by the National Science Foundation while working at
the Foundation. Any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not necessarily
reflect the views of the National Science Foundation. Choi's work is supported
in part by the National Science Foundation research grant DMS-1522687 and
a WSSSPE3 travel award. Hetherington was funded by
the Software Sustainability Institute, RCUK grants EP/H043160/1 and EP/N006410/1.

Work by Gunter was supported by the Office of Science, Office of Biological and 
Environmental Research, of the US Department of Energy under Award Numbers 
DE-AC02-05CH11231, DE-AC02-06CH11357, DE-AC05-00OR22725, and DE-AC02-98CH10886, 
as part of the DOE Systems Biology Knowledgebase and by the Office of Science, 
Office of Advanced Scientific Computing Research (ASCR) of the U.S. Department 
of Energy under Contract Number DE-AC02-05CH11231 as part of the Template
Interfaces for Agile Parallel Data-Intensive Science (TIGRES) project.

\todo{feel free to add stuff here}

>>>>

WSSSPE3 left the traditional Supercomputing Conference environment this year,
and in Turk's words, this spoke to the fact that scientific software comes from
many different types of inquiries, deployment, strategies for maintenance,
users, and ways of measuring the value of a piece of software. It appeared to
Turk that the supercomputing community generally adopts some top-down
approaches, whereas the SciPy community more often than not uses more bottom-up
systems. The essential messages perceived were also often bipolar: the
supercomputing community thinks that software is getting harder, with exascale
computing and optimization issues in mind; but the SciPy community thinks that
software is becoming better, with emerging tools such as Jupyter and
productivity packages for research workflows. Admitting such comparisons are
somewhat unfair generalizations, Turk reminded the audience that the different
approaches bring different types of ideas to the table, and he welcomed WSSSPE3
being conducted outside existing preconceptions.

@@

Turk reminded the audience that the volunteer model means unpaid labor. On this
note, he recommended Ashe Dryden's blog post on the ethics of unpaid labor and
the open source software
community\footnote{http://www.ashedryden.com/blog/the-ethics-of-unpaid-labor-and-the-oss-community}.
Oftentimes, a person funded to work full time on a scientific project can spend
a small amount of time for working on a piece of software necessary for that
project. However, researchers' abilities to participate in that volunteer
community are not always the same and may not always be aligned with their
research projects. From Turk's experience, we cannot always rely on unpaid labor
and volunteer time to sustain a piece of software---this came down to the
notions of the top-down and the bottom-up approaches, i.e., the funded versus
the grassroots. However, Turk pointed out that bottom-up, volunteer-driven
projects can be just as large-scale as a top-down software development project.

@@

Work by Katz was supported by the National Science Foundation while working at
the Foundation. Any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not necessarily
reflect the views of the National Science Foundation. Choi's work is supported
in part by the National Science Foundation research grant DMS-1522687 and
a WSSSPE3 travel award. Hetherington was funded by
the Software Sustainability Institute, RCUK grants EP/H043160/1 and EP/N006410/1.
%
Work by Gunter was supported by the Office of Science, Office of Biological and 
Environmental Research, of the US Department of Energy under Award Numbers 
DE-AC02-05CH11231, DE-AC02-06CH11357, DE-AC05-00OR22725, and DE-AC02-98CH10886, 
as part of the DOE Systems Biology Knowledgebase and by the Office of Science, 
Office of Advanced Scientific Computing Research (ASCR) of the U.S. Department 
of Energy under Contract Number DE-AC02-05CH11231 as part of the Template
Interfaces for Agile Parallel Data-Intensive Science (TIGRES) project.
%
\todo{feel free to add stuff here}
